A 2-Norm Condition Number for B\'{e}zier Curve Intersection

### 2019-04-07 doc/paper.tex
Putting paper in `elsarticle` format.
Also adding

- "Conclusion and Future Work" section
- "Acknowledgements" section
- Adding a "Condition Number in Practice" section name for the 3
  examples that were previously included in the section
  "Conditioning of Bezier Curve Intersection"
- Slight change in wording about "one of the four corners"

\documentclass[letterpaper,10pt]{article}

\usepackage[margin=1in]{geometry}

\usepackage{amsthm,amssymb,amsmath}
%% https://tex.stackexchange.com/a/396829/32270
\allowdisplaybreaks
\usepackage{bm}
%% H/T: https://tex.stackexchange.com/a/28334/32270
\usepackage{chngcntr}
\counterwithin{equation}{section}
%% H/T: https://tex.stackexchange.com/a/71153/32270
\usepackage[nottoc,notlot,notlof]{tocbibind}

@@

\usepackage[usenames, dvipsnames]{color}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  linkcolor=MidnightBlue,
  citecolor=ForestGreen,
  pdfinfo={
    CreationDate={D:20180816104534},
    ModDate={D:20180816104534},
  },
}

\usepackage{embedfile}
\embedfile{\jobname.tex}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{A 2-Norm Condition Number for B\'{e}zier Curve Intersection}
\rhead{Danny Hermes}

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\qed}{\(\blacksquare\)}
\newcommand{\reals}{\mathbf{R}}
\newcommand{\bigO}[1]{\mathcal{O}\left(#1\right)}
\newcommand{\eps}{\varepsilon}

\begin{document}

\begin{abstract}
\noindent We present a short note describing a condition number
of the intersection of two B\'{e}zier curves.
\\ \\
\noindent \emph{Keywords}: B\'{e}zier curve, Curve intersection,
Condition number, Numerical analysis
\end{abstract}

\tableofcontents

@@

\section{Conditioning of Generic Root-finding}

@@

Thus we know the maximum occurs at two of the four corners of the
rectangle. Due to sign cancellation, this leads to one of two values
\(\theta =
\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) \pm
2 \mu_1 \mu_2 \left(\bm{v} \cdot \bm{w}\right) +
\mu_2^2 \left(\bm{w} \cdot \bm{w}\right)\), the largest of which is
\(\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) +
2 \mu_1 \mu_2 \left|\bm{v} \cdot \bm{w}\right| +
\mu_2^2 \left(\bm{w} \cdot \bm{w}\right)\). Thus
\begin{equation}\label{eq:intersect-cond-num-too}
\kappa_{\alpha, \beta} = \sqrt{\frac{\mu_1^2
  \left(\bm{v} \cdot \bm{v}\right) +
  2 \mu_1 \mu_2 \left|\bm{v} \cdot \bm{w}\right| +
  \mu_2^2 \left(\bm{w} \cdot \bm{w}\right)}{\alpha^2 + \beta^2}}.
\end{equation}
Note that \(\widetilde{F}_{\pm} = \left[ \begin{array}{c} \mu_1 \\ \pm \mu_2
\end{array}\right]\), so
\begin{equation}
\left \lVert J\left(\alpha, \beta\right)^{-1} \widetilde{F}_{\pm}
  \right \rVert_2 = \left \lVert \left[ \begin{array}{c c} \bm{v} & \bm{w}
  \end{array}\right] \widetilde{F}_{\pm} \right \rVert_2 =
  \sqrt{\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) \pm
  2 \mu_1 \mu_2 \left(\bm{v} \cdot \bm{w}\right) +
  \mu_2^2 \left(\bm{w} \cdot \bm{w}\right)}
\end{equation}
gives each of the two values that produce the maximum value.
\end{proof}

@@

Consider the line
\(b_0(s) = \left[ \begin{array}{c} 2s \\ 2s \end{array}\right]\)
and improperly parameterized line
\(b_1(t) = \left[ \begin{array}{c} 4t^2 \\ 2 - 4t^2
\end{array}\right]\) which intersect at \(\alpha = \beta = 1/2\).
At the intersection we have \(J^{-1} = \frac{1}{8}
\left[ \begin{array}{c c} 2 & 2 \\ -1 & 1 \end{array}\right]\),
so that \(\bm{v} \cdot \bm{v} = \bm{w} \cdot \bm{w} =
5/64\) and \(\bm{v} \cdot \bm{w} = 3/64\). Since the
\(x\)-component of \(F(s, t)\) can be written as
\(2s - 4t^2 = 2 B_{1, 1}(s) - 4 B_{2, 2}(t)\) and the
\(y\)-component as \(2s + 4t^2 - 2 = 2 B_{1, 1}(s) - 2 B_{0, 2}(t)
- 2 B_{1, 2}(t) + 2 B_{2, 2}(t)\) we have
\begin{alignat}{2}
\mu_1 &= 2 B_{1, 1}(\alpha) &&+ 4 B_{2, 2}(\beta) = 2 \\
\mu_2 &= 2 B_{1, 1}(\alpha) + 2 B_{0, 2}(\beta) +
  2 B_{1, 2}(\beta) &&+ 2 B_{2, 2}(\beta) = 3.
\end{alignat}
Following~\eqref{eq:intersect-cond-num-too}, this gives
\(\kappa_{\alpha, \beta} = \sqrt{202}/8 \approx 1.78\).
This low condition number is somewhat unexpected since when using
the resultant to eliminate each parameter, one of the two roots is
a double root:
\begin{align}
\operatorname{Res}_t\left(x_0(s) - x_1(t), y_0(s) - y_1(t)\right) &=
  64(2s - 1)^2 \\
\operatorname{Res}_s\left(x_0(s) - x_1(t), y_0(s) - y_1(t)\right) &=
  4(2t - 1)(2t + 1).
\end{align}

@@

So, we see the condition number \(\kappa_{\alpha, \beta} = \sqrt{2}(2D + 1)\)
increases towards infinity as \(D\) does. This is what we expect as the
coefficients grow so large that their ratio approaches \(1\).

@@

\bibliography{paper}
\bibliographystyle{alpha}

>>>>

\documentclass[3p, authoryear, square]{elsarticle}
\bibliographystyle{elsarticle-harv}

\usepackage{amsthm,amssymb,amsmath}
%% https://tex.stackexchange.com/a/396829/32270
\allowdisplaybreaks
\usepackage{bm}
%% H/T: https://tex.stackexchange.com/a/28334/32270
\usepackage{chngcntr}
\counterwithin{equation}{section}
%% H/T: https://tex.stackexchange.com/a/26348/32270
\renewcommand*{\appendixname}{}
%% H/T: https://tex.stackexchange.com/a/23242/32270
\renewcommand*{\today}{April 7, 2019}

@@

\usepackage[usenames, dvipsnames]{color}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  pdfinfo={
    CreationDate={D:20180816104534},
    ModDate={D:20180816104534},
  },
}

\renewcommand{\qed}{\(\blacksquare\)}
\newcommand{\reals}{\mathbf{R}}
\newcommand{\bigO}[1]{\mathcal{O}\left(#1\right)}
\newcommand{\eps}{\varepsilon}

\journal{Computer Aided Geometric Design}

\begin{document}
%% H/T: https://tex.stackexchange.com/a/263503/32270
\hypersetup{
  urlcolor=MidnightBlue,
  linkcolor=MidnightBlue,
  citecolor=ForestGreen,
}

\begin{frontmatter}

\title{A 2-Norm Condition Number for B\'{e}zier Curve Intersection}
\author[djh]{Danny Hermes}\ead{dhermes@berkeley.edu}
\address[djh]{UC Berkeley, 970 Evans Hall \#3840, Berkeley, CA 94720-3840 USA}

\begin{abstract}
We present a short note describing a condition number
of the intersection of two B\'{e}zier curves.
\end{abstract}

\begin{keyword}
B\'{e}zier curve \sep Curve intersection \sep Condition number \sep
Numerical analysis
\end{keyword}

\end{frontmatter}

@@

\section{Conditioning of Generic Root-finding}\label{sec:generic}

@@

Thus we know the maximum occurs at one of the four corners of the
rectangle. Due to sign cancellation, this leads to one of two values
\(\theta =
\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) \pm
2 \mu_1 \mu_2 \left(\bm{v} \cdot \bm{w}\right) +
\mu_2^2 \left(\bm{w} \cdot \bm{w}\right)\), the largest of which is
\(\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) +
2 \mu_1 \mu_2 \left|\bm{v} \cdot \bm{w}\right| +
\mu_2^2 \left(\bm{w} \cdot \bm{w}\right)\). Thus
\begin{equation}\label{eq:intersect-cond-num-too}
\kappa_{\alpha, \beta} = \sqrt{\frac{\mu_1^2
  \left(\bm{v} \cdot \bm{v}\right) +
  2 \mu_1 \mu_2 \left|\bm{v} \cdot \bm{w}\right| +
  \mu_2^2 \left(\bm{w} \cdot \bm{w}\right)}{\alpha^2 + \beta^2}}.
\end{equation}
Note that \(\widetilde{F}_{\pm} = \left[ \begin{array}{c} \mu_1 \\ \pm \mu_2
\end{array}\right]\), so
\begin{equation}
\left \lVert J\left(\alpha, \beta\right)^{-1} \widetilde{F}_{\pm}
  \right \rVert_2 = \left \lVert \left[ \begin{array}{c c} \bm{v} & \bm{w}
  \end{array}\right] \widetilde{F}_{\pm} \right \rVert_2 =
  \sqrt{\mu_1^2 \left(\bm{v} \cdot \bm{v}\right) \pm
  2 \mu_1 \mu_2 \left(\bm{v} \cdot \bm{w}\right) +
  \mu_2^2 \left(\bm{w} \cdot \bm{w}\right)}
\end{equation}
gives each of the two values that could produce the maximum value.
\end{proof}

\section{Condition Number in Practice}

@@

Consider the line
\(b_0(s) = \left[ \begin{array}{c} 2s \\ 2s \end{array}\right]\)
and improperly parameterized line
\(b_1(t) = \left[ \begin{array}{c} 4t^2 \\ 2 - 4t^2
\end{array}\right]\) which intersect at \(\alpha = \beta = 1/2\).
At the intersection we have \(J^{-1} = \frac{1}{8}
\left[ \begin{array}{c c} 2 & 2 \\ -1 & 1 \end{array}\right]\),
so that \(\bm{v} \cdot \bm{v} = \bm{w} \cdot \bm{w} =
5/64\) and \(\bm{v} \cdot \bm{w} = 3/64\). Since the
\(x\)-component of \(F(s, t)\) can be written as
\(2s - 4t^2 = 2 B_{1, 1}(s) - 4 B_{2, 2}(t)\) and the
\(y\)-component as \(2s + 4t^2 - 2 = 2 B_{1, 1}(s) - 2 B_{0, 2}(t)
- 2 B_{1, 2}(t) + 2 B_{2, 2}(t)\) we have
\begin{alignat}{2}
\mu_1 &= 2 B_{1, 1}(\alpha) &&+ 4 B_{2, 2}(\beta) = 2 \\
\mu_2 &= 2 B_{1, 1}(\alpha) + 2 B_{0, 2}(\beta) +
  2 B_{1, 2}(\beta) &&+ 2 B_{2, 2}(\beta) = 3.
\end{alignat}
Following~\eqref{eq:intersect-cond-num-too}, this gives
\(\kappa_{\alpha, \beta} = \sqrt{202}/8 \approx 1.78\).
This low condition number is expected from a geometric point of view; i.e.
the intersection is a transversal intersection of two lines. However,
when using the resultant to eliminate each parameter, one of the two roots is
a double root:
\begin{align}
\operatorname{Res}_t\left(x_0(s) - x_1(t), y_0(s) - y_1(t)\right) &=
  64(2s - 1)^2 \\
\operatorname{Res}_s\left(x_0(s) - x_1(t), y_0(s) - y_1(t)\right) &=
  4(2t - 1)(2t + 1).
\end{align}
so an algebraic approach may lead to an incorrect conclusion that the
intersection is ill-conditioned.

@@

So, we see the condition number \(\kappa_{\alpha, \beta} = \sqrt{2}(2D + 1)\)
increases towards infinity as \(D\) does. This is what we expect as the
coefficients grow so large that their ratio \((1 + D) / D\) approaches \(1\).

@@

\section{Conclusion and Future Work}

The author hopes that this can be useful for evaluating and comparing the
performance of curve intersection implementations. By establishing a
straightforward and easy to compute closed form, the condition number can be
used more often to differentiate between cases where the algorithm or computer
code is at fault for loss in accuracy and cases where the conditioning of the
intersection itself is the cause. The framework set forth in
Section~\ref{sec:generic} can be applied in future work to compute the
condition number of higher order intersections such as surface-surface
intersections in \(\reals^3\).

\section{Acknowledgements}

The author would like to thank Professor Tom Sederberg for his useful
course notes (\cite{SederbergNotes}) that provide a useful introduction to
intersection methods as well as the survey \cite{Sederberg1986}. Additionally
\cite{Farouki1987} served as a great baseline when considering conditioning
of operations on B\'{e}zier Curves.

%% H/T: https://tex.stackexchange.com/a/137379/32270
\section*{\refname}
\bibliography{paper}
